---
title: "Script 2 - Spatio-temporal patterns of infection"
subtitle: "Blood parasites and avian health"
author: "Marie Buysse, Mathilde Ollagnier, Charly Souc, Marjorie Bruley, Thomas Blanchon, Carole Leray, Marion Vittecoq and Karen D. McCoy"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0.Librairies
```{r, message=FALSE, warning=FALSE}
library(readxl) #read_excel function
library(ggplot2)
library(dplyr) #filter function
library(lme4) #glmer function
library(MuMIn) #dredge function
library(kableExtra) #kable function
library(sjPlot) #tab_model function
library(DHARMa) #simulateResiduals function
library(mgcv) #gam function
library(gratia) #derivatives function
library(spdep) #for spatial auto-correlation analysis
library(sf) #st_as_sf and st_distance functions
library(spdep) #knn2nb function
```

# 1.Spatial auto-correlation of the probability of infection
Using a separate data file for each colony, year, and group (nests, then only chicks) (see `rawdata_spatial analysis` directpry), we first convert nest coordinates (x,y) to obtain approximate distances in meters between nest locations. We then use this data to create a spatial weights matrix for analyzing relationships between nest locations We define a neighbourhood size of 20 meters between nests and allowed for isolated nest locations. We expect infected clusters to occur at this spatial scale:
```
nests <- read.csv("carteau2023_juv.csv",sep = ";", dec = ".")
num_locations <- nrow(nests)
print (num_locations)
lat_to_meters <- function(lat) lat * 111000
lon_to_meters <- function(lon, lat) lon * 111000 * cos(lat * pi / 180)

nests$x_meters <- lon_to_meters(nests$x, nests$y)
nests$y_meters <- lat_to_meters(nests$y)

# Create spatial weights
coords <- cbind(nests$x_meters, nests$y_meters)
max_dist <- 20 # Adjust based on your study area
weights <- dnearneigh(coords, 0, max_dist)
weights_list <- nb2listw(weights, style="W", zero.policy = TRUE)

plot(coords, main="Spatial Distribution of Points")
```

We then examine the distances between nests by creating a distance matrix. We visualise these pairwise distances to validate the defined neighbourhood size:
```
nest_sf <- st_as_sf(nests, coords = c("x_meters", "y_meters"))
dist_matrix <- st_distance(nest_sf)

coords <- cbind(nests$x_meters, nests$y_meters)
knn1 <- knn2nb(knearneigh(coords, k=1))
max_nn_dist <- max(unlist(nbdists(knn1, coords)))

hist(as.vector(dist_matrix), breaks = 50, 
     main = "Distribution of Pairwise Distances Between Nests",
     xlab = "Distance (meters)")

summary(as.vector(dist_matrix))
```

We next convert the variable 'infected' into a categorical variable related to each location. This defines whether a location was positive or negative for Babesia infection. When considering the 'nid' datasets, a nest was positive if either the adult or a chick from the nest was infected. For the 'juv' datasets, we only considered whether chicks were infected or not.  This data is analysed for spatial autocorrelation using a Monte Carlo simulation of Join Count statistics, comparing the number of Infected-Infected, Uninfected-Uninfected, and Infected-Uninfected nest pairs to what would be expected by chance (null distribution from 999 simulations):
```
# Convert infected to a factor
nests$infected <- factor(nests$infected, levels = c(0, 1), labels = c("Uninfected", "Infected"))

# Optional: Monte Carlo version for more robust results
jc_mc <- joincount.mc(nests$infected, weights_list, nsim = 999, zero.policy = TRUE)
print(jc_mc)
str(jc_mc)
```

This step visualises the results:
```
# For Uninfected
ggplot(data.frame(sim = jc_mc[[1]]$res), aes(x = sim)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  geom_vline(xintercept = jc_mc[[1]]$statistic, color = "red", size = 1) +
  labs(title = "Simulated Join Counts (Uninfected)", 
       x = "Join Count", 
       y = "Frequency")

# For Infected
ggplot(data.frame(sim = jc_mc[[2]]$res), aes(x = sim)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  geom_vline(xintercept = jc_mc[[2]]$statistic, color = "red", size = 1) +
  labs(title = "Simulated Join Counts (Infected)", 
       x = "Join Count", 
       y = "Frequency")
```


# 2.Factors influencing the probability of infection
## 2.0. Dataset and prevalence data
```{r}
data_prev<- read_excel("Data_Spatio-temporal patterns of infection_infection_factors.xlsx")
data_prev$Site <- as.factor(data_prev$Site)
data_prev$Year <- as.factor(data_prev$Year)
data_prev$Age <- as.factor(data_prev$Age)
data_prev$Smear <- as.factor(data_prev$Smear)
data_prev$Sex <- as.factor(data_prev$Sex)
data_prev$qPCR <- as.factor(data_prev$qPCR)
data_prev$Nest <- as.factor(data_prev$Nest)
data_prev$Ticks_by_nest <- as.numeric(data_prev$Ticks_by_nest)
```
With:\
-Site: sampling colony (Carteau/Frioul)\
-Sample: ID of individual bird\
-Year: sampling year (2022/2023/2024)\
-Age: age of bird (Chick/Adult)\
-Smear: infection status based on blood smears technique (uninfected = 0/infected = 1)\
-Sex: sex of bird (F/M)\
-qPCR: infection status based on qPCR technique (uninfected = 0/infected = 1)\
-Nest: nest of bird\
-Ticks_by_nest: number of ticks collected in the nest\  

We compute the prevalence data for each detection dataset (`Smear` and `qPCR` according to `Age`, `Year` and `Site`).\
For the `Smear` dataset:
```{r, message=FALSE, warning=FALSE}
smear_dataset <- data_prev %>%
  filter(Smear %in% c(0,1))

prevalence_smear <- smear_dataset %>%
  group_by(Site, Year, Age) %>%
  summarise(
    n_total = n(),  
    n_positive = sum(Smear == 1, na.rm = TRUE),  
    prevalence = (n_positive / n_total) * 100) 

kable(prevalence_smear, align = "lllccc")
```

For the `qPCR` dataset:
```{r, message=FALSE, warning=FALSE}
qPCR_dataset <- data_prev %>%
  filter(qPCR %in% c(0,1))

prevalence_qPCR <- qPCR_dataset %>%
  group_by(Site, Year, Age) %>%
  summarise(
    n_total = n(),  
    n_positive = sum(qPCR == 1, na.rm = TRUE),  
    prevalence = (n_positive / n_total) * 100) 

kable(prevalence_qPCR, align = "lllccc")
```

We calculate the prevalence for all chicks and all adults:
```{r}
prevalence_qPCR_ad_ch <- qPCR_dataset %>%
  group_by(Age) %>%
  summarise(
    n_total = n(),  
    n_positive = sum(qPCR == 1, na.rm = TRUE),  
    prevalence = (n_positive / n_total) * 100,
    margin_error = 2 * sqrt((prevalence / 100) * (1 - (prevalence / 100)) / n_total) * 100)

kable(prevalence_qPCR_ad_ch, align = "lllcc")
```

We remove the missing values of variables that will be analysed in the model and create two data subsets (one for chicks, one for adults):
```{r}
data_model <- data_prev[!(is.na(data_prev$Nest)), ]
data_model <- data_model[!(is.na(data_model$qPCR)), ]
data_model <- data_model[!(is.na(data_model$Ticks_by_nest)), ]

data_chicks_model <- filter(data_model, Age == "Chick")
data_adults_model <- filter(data_model, Age == "Adult")
data_adults_model <- data_adults_model[!(is.na(data_adults_model$Sex)), ]
```


## 2.1.Factors influencing the probability of infection of adults
We include the variables `Sex`, `Year`, `Site`, and `Ticks_by_nest` as fixed effects and `Nest` as a random effect in our analyses based on GLMER model (binomial family). As the full model is complex, we first proceed by comparing null model to single-variable models:
```{r}
glmer_null_adults <-glmer(qPCR ~ 1 + (1|Nest), family=binomial(link = "logit"), data=data_adults_model)
glmer_year_adults <-glmer(qPCR ~ Year + (1|Nest), family=binomial(link = "logit"), data=data_adults_model)
glmer_site_adults <-glmer(qPCR ~ Site + (1|Nest), family=binomial(link = "logit"), data=data_adults_model)
glmer_ticks_adults <-glmer(qPCR ~ Ticks_by_nest + (1|Nest), family=binomial(link = "logit"), data=data_adults_model)
glmer_sex_adults <-glmer(qPCR ~ Sex + (1|Nest), family=binomial(link = "logit"), data=data_adults_model)

anova1 <- anova(glmer_null_adults, glmer_year_adults, test="Chisq")
anova2 <- anova(glmer_null_adults, glmer_site_adults, test="Chisq")
anova3 <- anova(glmer_null_adults, glmer_ticks_adults, test="Chisq")
anova4 <- anova(glmer_null_adults, glmer_sex_adults, test="Chisq")

p_value_anova1 <- anova1$`Pr(>Chisq)`[2]
p_value_anova2 <- anova2$`Pr(>Chisq)`[2]
p_value_anova3 <- anova3$`Pr(>Chisq)`[2]
p_value_anova4 <- anova4$`Pr(>Chisq)`[2]

anova_results_prev_ad_1 <- data.frame(
  Comparison = c("glmer_null",
                 "glmer_null vs Year",
                 "glmer_null vs Site",
                 "glmer_null vs Ticks_by_nest",
                 "glmer_null vs Sex"),
  ANOVA_p_value = c("-", p_value_anova1, p_value_anova2, p_value_anova3, p_value_anova4),
  AICc = c(AICc(glmer_null_adults), AICc(glmer_year_adults), AICc(glmer_site_adults), AICc(glmer_ticks_adults), AICc(glmer_sex_adults)))

# Print the table with kable
anova_results_prev_ad_1 %>%
  kable(caption = "Results of comparison between glmer_null_adults and single-variable models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
Based on these results, none of the ANOVA is significative and AICc of null model is the minimal, we select `qPCR ~ 1 + (1|Nest)` as the best-fitted model.\
**None of the included variables is a factor influencing the probability of infection of adults.**


## 2.2.Factors influencing the probability of infection of chicks
We include the variables `Year`, `Site`, and `Ticks_by_nest` as fixed effects and `Nest` as a random effect in our full GLMER model (binomial family):
```{r, warning=FALSE, message=FALSE}
glmer_prev_chicks_full <- glmer(qPCR ~ Year * Site * Ticks_by_nest + (1|Nest), family=binomial(link = "logit"), data=data_chicks_model, na.action = na.fail)
```

We use the `dredge` function to compare models:
```{r, message=FALSE, warning=FALSE}
model_dredge_chicks <- dredge(glmer_prev_chicks_full)
model_dredge_chicks
```

Based on `model_dredge_chicks`, the model with the minimal AICc value is `qPCR ~ Site + Ticks_by_nest + (1|Nest)`.\
To validate it as the best-fitted model, we compare it with all models having an AICc difference of 2 or less:
```{r}
glmer_prev_chicks_best <- glmer(qPCR ~ Site + Ticks_by_nest + (1|Nest), family=binomial(link = "logit"), data=data_chicks_model)
glmer_prev_chicks_year <- glmer(qPCR ~ Site + Ticks_by_nest + Year + (1|Nest), family=binomial(link = "logit"), data=data_chicks_model)
glmer_prev_chicks_int <- glmer(qPCR ~ Site * Ticks_by_nest + (1|Nest), family=binomial(link = "logit"), data=data_chicks_model)

anova2 <- anova(glmer_prev_chicks_best, glmer_prev_chicks_year, test="Chisq")
anova3 <- anova(glmer_prev_chicks_best, glmer_prev_chicks_int, test="Chisq")

p_value_anova2 <- anova2$`Pr(>Chisq)`[2]
p_value_anova3 <- anova3$`Pr(>Chisq)`[2]

anova_results_prev_ch <- data.frame(
  Comparison = c("glmer_prev_chicks_best",
                 "glmer_prev_chicks_year",
                 "glmer_prev_chicks_int",
                 "glmer_prev_chicks_best vs glmer_prev_chicks_year",
                 "glmer_prev_chicks_best vs glmer_prev_chicks_int"),
  ANOVA_p_value = c("-", "-", "-", p_value_anova2, p_value_anova3),
  AICc = c(AICc(glmer_prev_chicks_best), AICc(glmer_prev_chicks_year), AICc(glmer_prev_chicks_int),"-", "-"))

# Print the table with kable
anova_results_prev_ch %>%
  kable(caption = "Results of comparison between glmer_prev_chicks_best and other models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r,message=FALSE, warning=FALSE}
tab_model(glmer_prev_chicks_best, glmer_prev_chicks_year, glmer_prev_chicks_int, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

We examine the best-fitted model:
```{r}
summary(glmer_prev_chicks_best)
simulationOutput_chicks <- simulateResiduals(fittedModel = glmer_prev_chicks_best, plot = F)
plot(simulationOutput_chicks)
```

We select `qPCR ~ Site + Ticks_by_nest + (1|Nest)` as the best-fitted model.\
**The site and the exposure to ticks are factors influencing the probability of infection of chicks.**

# 3.Dynamics of infection in chicks
```{r, message=FALSE, warning=FALSE}
data_dyn <- read_excel("Data_Spatio-temporal patterns of infection_dynamics.xlsx")
data_dyn$Site <- as.factor(data_dyn$Site)
data_dyn$Sample <- as.factor(data_dyn$Sample)
data_dyn$Year <- as.factor(data_dyn$Year)
data_dyn$Age <- as.factor(data_dyn$Age)
data_dyn$Nest <- as.factor(data_dyn$Nest)
data_dyn$Sex <- as.factor(data_dyn$Sex)
data_dyn$Days <- as.numeric(data_dyn$Days)
data_dyn$Status_qPCR <- as.factor(data_dyn$Status_qPCR)
data_dyn$Ratio_qPCR <- as.numeric(data_dyn$Ratio_qPCR)
data_dyn$ErrorRatio_qPCR <- as.numeric(data_dyn$ErrorRatio_qPCR)
```
With:\
-Site: sampling colony (Carteau/Frioul)\
-Sample: ID of individual bird\
-Year: sampling year (2022/2023/2024)\
-Age: age of bird (Chick/Adult)\
-Nest: nest of bird\
-Sex: sex of bird (F/M)\
-Days: age in days of the chick at which the *Babesia* quantification was performed\
-Status_qPCR: infection status based on qPCR technique (uninfected/infected)\
-Ratio_qPCR: quantification of *Babesia* based on qPCR technique (ratio *Larus*/*Babesia* copies of genes)\
-ErrorRatio_qPCR: ratio error associated with `Ratio_qPCR`\

**Important information:**\
This dataset includes only chicks monitored during 2024 because these individuals were monitored through several blood samples in order to perform *Babesia* quantification using qPCR technique.\

We remove the missing values of variables that will be analysed in the model and we log-transform the values of `Ratio_qPCR`:
```{r}
data_dyn <- data_dyn[complete.cases(data_dyn[ , 10]),] #based on missing Status_qPCR
data_ratio <- data_dyn[complete.cases(data_dyn[ , 11]),] #based on missing Ratio_qPCR
data_ratio_no0 <- data_ratio[-which(data_ratio$Status_qPCR == "uninfected"),]

data_ratio_no0$log_ratioqPCR <- log10(data_ratio_no0$Ratio_qPCR)
```

## 3.1.GAM model
We include the variable `Days` (estimated age of the chick at the time of the infection) as a smooth term and the variables `Site` and `Sample` as random effects in our full GAM model:
```{r}
gam_model <- gam(log_ratioqPCR ~ s(Days) + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)
gam_model_null <- gam(log_ratioqPCR ~ 1 + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)
anova(gam_model, gam_model_null, test = "Chisq")
AIC(gam_model, gam_model_null)
```

Based on ANOVA result and AICc values, **the best-fitted model is `log_ratioqPCR ~ s(Days) + s(Sample, bs = "re") + s(Site, bs = "re")`**.\  

We examine this model with different functions:
```{r, message=FALSE, warning=FALSE}
summary(gam_model)
plot(gam_model)

simulationOutput_gam <- simulateResiduals(fittedModel = gam_model)
plot(simulationOutput_gam)
```

**The parasitemia changes over the growth of chicks (p-value = 0.000628 with R-sq.(adj) = 0.493).**\  

To develop this analysis, we identify at which day the parasitemia is decreasing based on the derivative values:
```{r, message=FALSE, warning=FALSE}
derivs <- derivatives(gam_model, term = "s(Days)", type = "central")
draw(derivatives(gam_model, term = "s(Days)"))
```

**According to the table `derivs`, the derivative becomes negative at day 37.**\  

## 3.2.Visualisation
Here is a visualisation of the data with the model:
```{r, warning=FALSE}
# Generate new data for predictions
gam_data <- data.frame(
  Days = seq(min(data_ratio_no0$Days), max(data_ratio_no0$Days), length.out = 100),
  Sample = levels(data_ratio_no0$Sample)[1], # Example: First level of Sample
  Site = levels(data_ratio_no0$Site)[1]      # Example: First level of Site
)

# Predictions including random effects, with standard errors
predictions <- predict(gam_model, newdata = gam_data, type = "link", se.fit = TRUE, re.form = NULL)

# Add predictions and confidence intervals to the dataframe
gam_data$fit <- predictions$fit                      # Predicted values
gam_data$se.fit <- predictions$se.fit                # Standard errors
gam_data$lower <- gam_data$fit - 1.96 * gam_data$se.fit  # Lower bound of CI
gam_data$upper <- gam_data$fit + 1.96 * gam_data$se.fit  # Upper bound of CI

ggplot() +
  geom_point(data = data_ratio_no0, aes(x = Days, y = log_ratioqPCR), size = 2) +
  geom_line(data = gam_data, aes(x = Days, y = fit), color = "blue", size = 1) +
  geom_ribbon(data = gam_data, aes(x = Days, ymin = lower, ymax = upper), alpha = 0.2, fill = "blue") +
  theme_minimal() +
  labs(x = "Days", y = "log_ratioqPCR", title = "Parasitemia over time")
```

## 3.3.Complementary test to assess a potential effect of Sex
We additionally examine if the variable `Sex` could impact the model: 
```{r}
gam_model_full <- gam(log_ratioqPCR ~ s(Days) + Sex + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)
gam_model_sex <- gam(log_ratioqPCR ~ Sex + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)
gam_model_days <- gam(log_ratioqPCR ~ s(Days) + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)
gam_model_null <- gam(log_ratioqPCR ~ 1 + s(Sample, bs = "re") + s(Site, bs = "re"), data = data_ratio_no0)

AIC(gam_model_full, gam_model_sex, gam_model_days, gam_model_null)

anova(gam_model_full, gam_model_days, test = "Chisq")
```
Even if AICc values are close between model full (139.9858) and model days (139.2792), there is no significant difference based on ANOVA.\  

**`Sex` is not a significant fixed effect to explain parasitemia over time.**