---
title: "Script 3 - Impacts on individual physiology"
subtitle: "Blood parasites and avian health"
author: "Marie Buysse, Mathilde Ollagnier, Charly Souc, Marjorie Bruley, Thomas Blanchon, Carole Leray, Marion Vittecoq and Karen D. McCoy"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0.Librairies
```{r, message=FALSE, warning=FALSE}
library(readxl) #read_excel function
library(ggplot2)
library(dplyr)
library(kableExtra) #kable function
library(ggpubr) #ggarrange function
library(lme4) #lmer function
library(MuMIn) #dredge function
library(sjPlot) #tab_model function
library(DHARMa) #simulateResiduals function
library(lmtest) #bptest, dwtest functions
```

**Information about data visualisation:**\
For a same variable, one data visualisation is based on fixed effect selected in the best-fitted model and most significant random effects. Another visualisation include the infection status or parasitemia, whether significant or not.

# 1. Data
## 1.1. Data loading
```{r}
data <- read_excel("Data_Impacts on individual physiology.xlsx")
data$Site <- as.factor(data$Site)
data$Sample <- as.factor(data$Sample)
data$Year <- as.factor(data$Year)
data$Age <- as.factor(data$Age)
data$Status_qPCR <- as.factor(data$Status_qPCR)
data$Ratio_qPCR <- as.numeric(data$Ratio_qPCR)
data$ErrorRatio_qPCR <- as.numeric(data$ErrorRatio_qPCR)
data$Hemoglobin <- as.numeric(data$Hemoglobin)
data$Weight <- as.numeric(data$Weight)
data$Tarsus <- as.numeric(data$Tarsus)
data$RatioWT <- as.numeric(data$RatioWT)
data$Sex <- as.factor(data$Sex)
```
With:\
-Site: sampling colony (Carteau/Frioul)\
-Sample: ID of individual bird\
-Year: sampling year (2022/2023/2024)\
-Age: age of bird (Chick/Adult)\
-Status_qPCR: infection status based on qPCR technique (uninfected/infected)\
-Ratio_qPCR: quantification of *Babesia* based on qPCR technique (ratio *Larus*/*Babesia* copies of genes)\
-ErrorRatio_qPCR: ratio error associated with `Ratio_qPCR`\
-Hemoglobin: measure of circulating concentration of hemoglobin\
-Tarsus: length of the tarsus\
-Weight: bird weight\
-RatioWT: ratio between `Weight` and `Tarsus`\
-Sex: sex of bird (F/M)\  


## 1.2. Preliminary analyses
### 1.2.1. Metrics about parasitemia
```{r}
data_parasitemia <- data[!(is.na(data$Ratio_qPCR)), ]
```

```{r, message=FALSE, warning=FALSE}
summary_table_parasitemia <- data_parasitemia %>%
  group_by(Age) %>%
  summarise(
    Mean = mean(Ratio_qPCR),
    SD = sd(Ratio_qPCR),
    SE = sd(Ratio_qPCR, na.rm = TRUE) / sqrt(length(Ratio_qPCR)),
    N = n())

summary_table_parasitemia %>%
  kable(caption = "Metrics about Ratio_qPCR based on Age") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### 1.2.2. Circulating concentration of hemoglobin
Let's see first if `Hemoglobin` levels are equivalent between chicks and adults:
```{r}
a <- data$Hemoglobin[data$Age=="Chick"]
b <- data$Hemoglobin[data$Age=="Adult"]
wilcox.test(a,b, paired=FALSE)
```

Since p-value is much smaller than 0.05 (< 2.2e-16), the `Hemoglobin` levels significantly differ between chicks and adults.\
**`Hemoglobin` analyses will be conducted on separate datasets based on `Age`.**

```{r}
data_hemo <- data[!(is.na(data$Hemoglobin)), ]
data_hemo_adults <- filter(data_hemo, Age == "Adult")
data_hemo_chicks <- filter(data_hemo, Age == "Chick")
```

Here are some metrics about `Hemoglobin` for each dataset:
```{r, message=FALSE, warning=FALSE}
summary_table_1 <- data_hemo %>%
  group_by(Age) %>%
  summarise(
    Mean = mean(Hemoglobin),
    SD = sd(Hemoglobin),
    SE = sd(Hemoglobin, na.rm = TRUE) / sqrt(length(Hemoglobin)),
    N = n(),
    CI_lower = Mean - qt(0.975, df = N - 1) * (SD / sqrt(N)),
    CI_upper = Mean + qt(0.975, df = N - 1) * (SD / sqrt(N))
  )

summary_table_1 %>%
  kable(caption = "Metrics about Hemoglobin based on Age") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


```{r, message=FALSE, warning=FALSE}
data_hemo_stats <- data_hemo[!(is.na(data_hemo$Status_qPCR)), ]
summary_table_2 <- data_hemo_stats %>%
  group_by(Age, Status_qPCR) %>%
  summarise(
    Mean = mean(Hemoglobin),
    SD = sd(Hemoglobin),
    SE = sd(Hemoglobin, na.rm = TRUE) / sqrt(length(Hemoglobin)),
    N = n(),
    CI_lower = Mean - qt(0.975, df = N - 1) * (SD / sqrt(N)),
    CI_upper = Mean + qt(0.975, df = N - 1) * (SD / sqrt(N))
  )

summary_table_2 %>%
  kable(caption = "Metrics about Hemoglobin based on Age and infection status") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### 1.2.2.1. Data distribution of `Hemoglobin` in adults
In the dataset `data_hemo_adults`:
```{r, message=FALSE, warning=FALSE}
ggplot(data_hemo_adults, aes(x=Hemoglobin)) + geom_histogram()
shapiro.test(data_hemo_adults$Hemoglobin)
```
Since p-value is greater than 0.05 (p-value = 0.2172), **`data_hemo_adults$Hemoblobin` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\  

In the dataset `data_hemo_adults` with only individuals with `Ratio_qPCR`:
```{r}
data_hemo_ad_ratioBab <- data_hemo_adults[!(is.na(data_hemo_adults$Ratio_qPCR)), ]
```

Considering that this dataset includes only 10 individuals, **the impact of the parasitemia (`Ratio_qPCR`) on `Hemoglobin` in adults will not be examined by models but rather by simple comparison tests**.\ 


### 1.2.2.2. Data distribution of `Hemoglobin` in chicks
```{r, message=FALSE, warning=FALSE}
ggplot(data_hemo_chicks, aes(x=Hemoglobin)) + geom_histogram()
shapiro.test(data_hemo_chicks$Hemoglobin)
```
Since p-value is greater than 0.05 (p-value = 0.6399), **`data_hemo_chicks$Hemoblobin` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\  

In the dataset `data_hemo_chicks` with only individuals with `Ratio_qPCR`:
```{r}
data_hemo_ch_ratioBab <- data_hemo_chicks[!(is.na(data_hemo_chicks$Ratio_qPCR)), ]
data_hemo_ch_ratioBab <- data_hemo_ch_ratioBab[!(is.na(data_hemo_ch_ratioBab$ErrorRatio_qPCR)), ]

ggplot(data_hemo_ch_ratioBab, aes(x=Hemoglobin)) + geom_histogram()
shapiro.test(data_hemo_ch_ratioBab$Hemoglobin)
```
Since p-value is greater than 0.05 (p-value = 0.9963), **`data_hemo_ch_ratioBab$Hemoblobin` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\  


### 1.2.3. Body condition
**Body condition is considered as `RatioWT`.**\  

Let's see first if `RatioWT` levels are equivalent between chicks and adults:
```{r}
c <- data$RatioWT[data$Age=="Chick"]
d <- data$RatioWT[data$Age=="Adult"]
wilcox.test(c,d, paired=FALSE)
```

Since p-value is much smaller than 0.05 (< 2.2e-16), the `RatioWT` levels significantly differ between chicks and adults.\
**`RatioWT` analyses will be conducted on separate datasets based on `Age`.**

```{r}
data_ratiowt <- data[!(is.na(data$RatioWT)), ]
data_ratiowt_adults <- filter(data_ratiowt, Age == "Adult")
data_ratiowt_chicks <- filter(data_ratiowt, Age == "Chick")
```

Here are some metrics about `RatioWT` for each dataset:
```{r, message=FALSE, warning=FALSE}
summary_table_1 <- data_ratiowt %>%
  group_by(Age) %>%
  summarise(
    Mean = mean(RatioWT),
    SD = sd(RatioWT),
    SE = sd(RatioWT, na.rm = TRUE) / sqrt(length(RatioWT)),
    N = n(),
    CI_lower = Mean - qt(0.975, df = N - 1) * (SD / sqrt(N)),
    CI_upper = Mean + qt(0.975, df = N - 1) * (SD / sqrt(N))
  )

summary_table_1 %>%
  kable(caption = "Metrics about RatioWT based on Age") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


```{r, message=FALSE, warning=FALSE}
data_ratiowt_stats <- data_ratiowt[!(is.na(data_ratiowt$Status_qPCR)), ]
summary_table_2 <- data_ratiowt_stats %>%
  group_by(Age, Status_qPCR) %>%
  summarise(
    Mean = mean(RatioWT),
    SD = sd(RatioWT),
    SE = sd(RatioWT, na.rm = TRUE) / sqrt(length(RatioWT)),
    N = n(),
    CI_lower = Mean - qt(0.975, df = N - 1) * (SD / sqrt(N)),
    CI_upper = Mean + qt(0.975, df = N - 1) * (SD / sqrt(N))
  )

summary_table_2 %>%
  kable(caption = "Metrics about RatioWT based on Age and infection status") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### 1.2.3.1. Data distribution of `RatioWT` in adults 
In the dataset `data_ratiowt_adults`:
```{r, message=FALSE, warning=FALSE}
ggplot(data_ratiowt_adults, aes(x=RatioWT)) + geom_histogram()
shapiro.test(data_ratiowt_adults$RatioWT)
```
Since p-value is greater than 0.05 (p-value = 0.7556), **`data_ratiowt_adults$RatioWT` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\

In the dataset `data_ratiowt_adults` with only individuals with `Ratio_qPCR`:
```{r}
data_ratiowt_ad_ratioBab <- data_ratiowt_adults[!(is.na(data_ratiowt_adults$Ratio_qPCR)), ]
```

Considering that this dataset includes only 10 individuals, **the impact of the parasitemia (`Ratio_qPCR`) on `RatioWT` in adults will not be examined by models but rather by simple comparison tests**.\   


### 1.2.3.2. Data distribution of `RatioWT` in chicks
In the dataset `data_ratiowt_chicks`:
```{r, message=FALSE, warning=FALSE}
e <- ggplot(data_ratiowt_chicks, aes(x=RatioWT)) + geom_histogram()
shap_test1 <- shapiro.test(data_ratiowt_chicks$RatioWT)

data_ratiowt_chicks$log_ratiowt <- log10(data_ratiowt_chicks$RatioWT)
f <- ggplot(data_ratiowt_chicks, aes(x=log_ratiowt)) + geom_histogram()
shap_test2 <- shapiro.test(data_ratiowt_chicks$log_ratiowt)

ggarrange(e,f, labels = c("RatioWT in data_ratiowt_chicks", "log10(RatioWT) in data_ratiowt_chicks"), ncol = 1, nrow = 2)

shapiro_results_hemo <- data.frame(
  Tests = c("Shapiro test on RatioWT in chicks", "Shapiro test on log10(RatioWT) in chicks"),
  p_value = c(shap_test1$p.value, shap_test2$p.value))

shapiro_results_hemo %>%
  kable(caption = "Shapiro test results for RatioWT in data_ratiowt_chicks") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Since p-value is greater than 0.05 (p-value = 0.4810149), **`data_ratiowt_chicks$log_ratiowt` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\  

In the dataset `data_ratiowt_chicks` with only individuals with `Ratio_qPCR`:
```{r, message=FALSE, warning=FALSE}
data_ratiowt_ch_ratioBab <- data_ratiowt_chicks[!(is.na(data_ratiowt_chicks$Ratio_qPCR)), ]
data_ratiowt_ch_ratioBab <- data_ratiowt_ch_ratioBab[!(is.na(data_ratiowt_ch_ratioBab$ErrorRatio_qPCR)), ]

ggplot(data_ratiowt_ch_ratioBab, aes(x=RatioWT)) + geom_histogram()
shapiro.test(data_ratiowt_ch_ratioBab$RatioWT)
```
Since p-value is greater than 0.05 (p-value = 0.3707), **`data_ratiowt_ch_ratioBab$RatioWT` variable is normally distributed**. We will perform the analyses using LMER and check for distribution of model residuals.\  


# 2. Effects of infection on circulating concentration of hemoglobin
## 2.1. Based on infectious status (binary: infected/uninfected)
### 2.1.1. In adults
We remove the missing values of variables that will be analysed in the model:
```{r}
data_hemo_ad_model <- data_hemo_adults[!(is.na(data_hemo_adults$Status_qPCR)), ]
data_hemo_ad_model <- data_hemo_ad_model[!(is.na(data_hemo_ad_model$Sex)), ]
```

We include the variables `Status_qPCR` and `RatioWT` as fixed effects and `Site`, `Sex`, and `Year` as random effects in our full LMER model:
```{r, message=FALSE, warning=FALSE}
lmer_hemo_ad_full <- lmer(Hemoglobin ~ Status_qPCR * RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ad_model, na.action = na.fail)
```

We use the dredge function to compare models:
```{r, message=FALSE, warning=FALSE}
model_dredge_hemo_ad <- dredge(lmer_hemo_ad_full)
model_dredge_hemo_ad
```

Based on `model_dredge_hemo_ad`, the model with the minimal AICc value is `RatioWT + (1|Site) + (1|Sex) + (1|Year)`.\
To validate it as the best-fitted model, we compare it with all models having an AICc difference of 2 or less:
```{r message=FALSE, warning=FALSE}
lmer_hemo_ad_1 <- lmer(Hemoglobin ~ RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ad_model, na.action = na.fail)
lmer_hemo_ad_2 <- lmer(Hemoglobin ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ad_model, na.action = na.fail)
lmer_hemo_ad_3 <- lmer(Hemoglobin ~ RatioWT + Status_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ad_model, na.action = na.fail)
lmer_hemo_ad_4 <- lmer(Hemoglobin ~ Status_qPCR * RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ad_model, na.action = na.fail)

anova1 <- anova(lmer_hemo_ad_1, lmer_hemo_ad_2, test="Chisq")
anova2 <- anova(lmer_hemo_ad_1, lmer_hemo_ad_3, test="Chisq")
anova3 <- anova(lmer_hemo_ad_1, lmer_hemo_ad_4, test="Chisq")

p_value_anova1 <- anova1$`Pr(>Chisq)`[2]
p_value_anova2 <- anova2$`Pr(>Chisq)`[2]
p_value_anova3 <- anova3$`Pr(>Chisq)`[2]

anova_results_hemo_ad <- data.frame(
  Comparison = c("RatioWT (best)",
                 "best vs null",
                 "best vs RatioWT + Status_qPCR",
                 "best vs full"),
  ANOVA_p_value = c("-", p_value_anova1, p_value_anova2, p_value_anova3),
  AICc = c(AICc(lmer_hemo_ad_1), AICc(lmer_hemo_ad_2), AICc(lmer_hemo_ad_3),AICc(lmer_hemo_ad_4)))

# Print the table with kable
anova_results_hemo_ad %>%
  kable(caption = "Results of comparison between best AICc model and other models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r}
tab_model(lmer_hemo_ad_1, lmer_hemo_ad_2, lmer_hemo_ad_3, lmer_hemo_ad_4, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results, we select `RatioWT + (1|Site) + (1|Sex) + (1|Year)` as the best-fitted model.\

We examine the best-fitted model:
```{r}
summary(lmer_hemo_ad_1)

simulationOutput_hemo_ad <- simulateResiduals(fittedModel = lmer_hemo_ad_1, plot = F)
plot(simulationOutput_hemo_ad)

shapiro.test(residuals(lmer_hemo_ad_1))
```

**The ratio between mass/tarsus (body condition) is a fixed effect influencing the circulating concentration of hemoglobin in adults, but the presence of *Babesia* does not appear to have any impact.**\  

Visualisation: 
```{r}
plotA <- ggplot(data_hemo_ad_model, aes(x= RatioWT, y= Hemoglobin, color=Sex, fill=Sex))+
  geom_point() +
  labs(x = "Ratio mass/tarsus (body condition)",
       y = "Concentration of hemoglobin", 
       title = "Hemoglobin in adults according to sex and body condition")+ 
  theme(plot.title = element_text(size = 9)) + 
  facet_grid(~Sex) + 
  theme(legend.position = "none")

plotB <- ggplot(data_hemo_ad_model, aes(x= Status_qPCR, y= Hemoglobin, color=Status_qPCR, fill=Status_qPCR))+
  geom_point() +
  geom_boxplot(alpha= 0.5) + 
  labs(x = "Infection status",
       y = "Concentration of hemoglobin", 
       title = "Hemoglobin in adults according to sex and infection status")+ 
  theme(plot.title = element_text(size = 9), 
        axis.text.x = element_blank()) + 
  facet_grid(~Sex)

ggarrange(plotA, plotB, nrow = 1, ncol = 2)
```


### 2.1.2. In chicks
We remove the missing values of variables that will be analysed in the model:
```{r}
data_hemo_ch_model <- data_hemo_chicks[!(is.na(data_hemo_chicks$Status_qPCR)), ]
data_hemo_ch_model <- data_hemo_ch_model[!(is.na(data_hemo_ch_model$Sex)), ]
```

We include the variables `Status_qPCR` and `RatioWT` as fixed effects and `Site`, `Sex`, and `Year` as random effects in our full LMER model:
```{r, message=FALSE, warning=FALSE}
lmer_hemo_ch_full <- lmer(Hemoglobin ~ Status_qPCR * RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_model, na.action = na.fail)
```

We use the dredge function to compare models:
```{r, message=FALSE, warning=FALSE}
model_dredge_hemo_ch <- dredge(lmer_hemo_ch_full)
model_dredge_hemo_ch
```

Based on `model_dredge_hemo_ch`, the model with the minimal AICc value is `1 + (1|Site) + (1|Sex) + (1|Year)`.\
To validate it as the best-fitted model, we compare it with all models having an AICc difference of 2 or less:
```{r message=FALSE, warning=FALSE}
lmer_hemo_ch_1 <- lmer(Hemoglobin ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_model, na.action = na.fail)
lmer_hemo_ch_2 <- lmer(Hemoglobin ~ Status_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_model, na.action = na.fail)

anova4 <- anova(lmer_hemo_ch_1, lmer_hemo_ch_2, test="Chisq")

p_value_anova4 <- anova4$`Pr(>Chisq)`[2]

anova_results_hemo_ch <- data.frame(
  Comparison = c("null (best)",
                 "best vs Status_qPCR"),
  ANOVA_p_value = c("-", p_value_anova4),
  AICc = c(AICc(lmer_hemo_ch_1), AICc(lmer_hemo_ch_2)))

# Print the table with kable
anova_results_hemo_ch %>%
  kable(caption = "Results of comparison between best AICc model and other models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r}
tab_model(lmer_hemo_ch_1, lmer_hemo_ch_2, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results, we select `1 + (1|Site) + (1|Sex) + (1|Year)` as the best-fitted model.
```{r}
shapiro.test(residuals(lmer_hemo_ch_1))
```

**Based on the measured variables, we did not identified a fixed effect influencing the circulating concentration of hemoglobin in chicks, including the presence of *Babesia*.**

Visualisation: 
```{r}
plotC <- ggplot(data_hemo_ch_model, aes(x= Sex, y= Hemoglobin, color=Sex, fill=Sex))+
  geom_point() +
  geom_boxplot(alpha=0.5) +
  labs(x = "Sex",
       y = "Concentration of hemoglobin", 
       title = "Hemoglobin in chicks according to sex and site") + 
  theme(axis.text.x = element_blank(),
        plot.title = element_text(size = 9)) + 
  facet_grid(~Site)

plotD <- ggplot(data_hemo_ch_model, aes(x= Status_qPCR, y= Hemoglobin, color=Status_qPCR, fill=Status_qPCR))+
  geom_point() +
  geom_boxplot(alpha=0.5) +
  labs(x = "Infection status",
       y = "Concentration of hemoglobin", 
       title = "Hemoglobin in chicks according to sex, site, and infection status") + 
  theme(axis.text.x = element_blank(),
        plot.title = element_text(size = 9)) + 
  facet_grid(Site~Sex)

ggarrange(plotC, plotD, nrow = 1, ncol = 2)
```


## 2.2. Based on parasitemia (continuous)
### 2.2.1. In adults
To determine whether a Pearson Correlation test can be used, the data must be tested for normality, autocorrelation and homoscedasticity:
```{r, message=FALSE, warning=FALSE}
shapiro.test(data_hemo_ad_ratioBab$Hemoglobin)
```
**p-value > 0.05, the data is normally distributed.**\

```{r, message=FALSE, warning=FALSE}
bptest(aov(Hemoglobin ~ Ratio_qPCR, data = data_hemo_ad_ratioBab)) 
```
**p-value > 0.05, the null hypothesis of residual homoscedasticity cannot be rejected.**\

```{r, message=FALSE, warning=FALSE}
dwtest(aov(Hemoglobin ~ Ratio_qPCR, data = data_hemo_ad_ratioBab))
```
**p-value > 0.05, the null hypothesis that the residuals are not autocorrelated cannot be rejected.**\

We can use a Pearson Correlation test: 
```{r, message=FALSE, warning=FALSE}
hemo_adults <- data_hemo_ad_ratioBab$Hemoglobin
ratioqPCR_adults <- data_hemo_ad_ratioBab$Ratio_qPCR
cor.test(hemo_adults,ratioqPCR_adults, method="pearson")
```

**p-value > 0.05 (p-value = 0.3821 & cor = 0.3107673), the null hypothesis that there is no correlation between the two variables `Hemoglobin` and `Ratio_qPCR` cannot be rejected.**\
**The parasitemia doesn't influence the circulating concentration of hemoglobin in adults.**


### 2.2.2. In chicks
We remove the missing values of variables that will be analysed in the model:
```{r}
data_hemo_ch_model <- data_hemo_chicks[!(is.na(data_hemo_chicks$Status_qPCR)), ]
data_hemo_ch_model <- data_hemo_ch_model[!(is.na(data_hemo_ch_model$Sex)), ]
data_hemo_ch_ratioBab <- data_hemo_ch_model[!(is.na(data_hemo_ch_model$Ratio_qPCR)), ]
data_hemo_ch_ratioBab <- data_hemo_ch_ratioBab[!(is.na(data_hemo_ch_ratioBab$ErrorRatio_qPCR)), ]
```

We include the variables `Ratio_qPCR` and `RatioWT` as fixed effects and `Site`, `Sex`, and `Year` as random effects in our full LMER model. We will analyse `Ratio_qPCR` using a log-transformation:
```{r, message=FALSE, warning=FALSE}
data_hemo_ch_ratioBab$log_ratio_qPCR <- log10(data_hemo_ch_ratioBab$Ratio_qPCR)

lmer_hemo_ch_ratio_full <- lmer(Hemoglobin ~ log_ratio_qPCR * RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_ratioBab, na.action = na.fail)
```

We use the dredge function to compare models:
```{r, message=FALSE, warning=FALSE}
model_dredge_hemo_ratio_ch <- dredge(lmer_hemo_ch_ratio_full)
model_dredge_hemo_ratio_ch
```

Based on these results, the model with the lower AICc is `1 + (1|Site) + (1|Sex) + (1|Year)`.\
To validate it, we will also compare it with the models with AICc difference of 2 and including a comparison between the null against `log_ratio_qPCR`:
```{r, message=FALSE, warning=FALSE}
lmer_hemo_ch_ratio_null <- lmer(Hemoglobin ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_ratioBab, na.action = na.fail)
lmer_hemo_ch_ratio_1 <- lmer(Hemoglobin ~ RatioWT + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_ratioBab, na.action = na.fail)
lmer_hemo_ch_ratio_2 <- lmer(Hemoglobin ~ log_ratio_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_hemo_ch_ratioBab, na.action = na.fail)

anova(lmer_hemo_ch_ratio_null, lmer_hemo_ch_ratio_1, test="Chisq")
anova(lmer_hemo_ch_ratio_null, lmer_hemo_ch_ratio_2, test="Chisq")
```

We examine the models (NB: DHARMA cannot be used on weighted LMER):
```{r, message=FALSE, warning=FALSE}
tab_model(lmer_hemo_ch_ratio_null, lmer_hemo_ch_ratio_1, lmer_hemo_ch_ratio_2, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results:\
-The AICc of the `RatioWT` model is not the lowest but the ANOVA indicates an effect of the adding of `RatioWT`.\
-The model with `log_ratio_qPCR` alone suggests a negative correlation with `Hemoglobin` but this correlation isn't significant. A further investigation and bigger dataset would help to study this correlation.\

We choose the model `RatioWT + (1|Site) + (1|Sex) + (1|Year)` as the best-fitted model.\
We examine the best-fitted model:
```{r}
summary(lmer_hemo_ch_ratio_1)
plot(lmer_hemo_ch_ratio_1)
qqnorm(resid(lmer_hemo_ch_ratio_1)); qqline(resid(lmer_hemo_ch_ratio_1))
shapiro.test(residuals(lmer_hemo_ch_ratio_1))
```

**Conclusion:**\
**The parasitemia might have an effect of concentration of hemoglobin in chicks with a bigger dataset but here there is no effect of the parasitemia on the concentration of hemoglobin.**\  

Visualisation: 
```{r,message=FALSE, warning=FALSE}
plotE <- ggplot(data_hemo_ch_ratioBab, aes(x = RatioWT, y = Hemoglobin)) +
  geom_point() +  
   geom_smooth(method = lm) +  
  labs(x = "Body condition",
       y = "Concentration of hemoglobin", 
       title = "Concentration of hemoglobin in chicks according to body condition") +  
  theme(plot.title = element_text(size = 9))

plotF <- ggplot(data_hemo_ch_ratioBab, aes(x = log_ratio_qPCR, y = Hemoglobin)) +
  geom_point() +  
  labs(x = "log10(RatioqPCR)",
       y = "Concentration of hemoglobin", 
       title = "Concentration of hemoglobin in chicks according to parasitemia") +  
  theme(plot.title = element_text(size = 9))

ggarrange(plotE, plotF, nrow = 1, ncol = 2)
```


# 3. Effects of infection on body condition
## 3.1. Based on infectious status (binary: infected/uninfected) 
### 3.1.1. In adults
We remove the missing values of variables that will be analysed in the model:
```{r}
data_ratiowt_ad_model <- data_ratiowt_adults[!(is.na(data_ratiowt_adults$Status_qPCR)), ]
data_ratiowt_ad_model <- data_ratiowt_ad_model[!(is.na(data_ratiowt_ad_model$Sex)), ]
```

We include the variables `Status_qPCR` as a fixed effect and `Site`, `Sex`, and `Year` as random effects in our full LMER model: 
```{r, message=FALSE, warning=FALSE}
lmer_ratiowt_ad_full <- lmer(RatioWT ~ Status_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ad_model)
```

As the model is rather simple, we will directly use ANOVA and AICc comparison:
```{r, message=FALSE, warning=FALSE}
lmer_ratiowt_ad_null <- lmer(RatioWT ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ad_model)

anova5 <- anova(lmer_ratiowt_ad_full, lmer_ratiowt_ad_null, test="Chisq")

p_value_anova5 <- anova5$`Pr(>Chisq)`[2]

anova_results_ratiowt_ad <- data.frame(
  Comparison = c("Full",
                 "Full vs null"),
  ANOVA_p_value = c("-", p_value_anova5),
  AICc = c(AICc(lmer_ratiowt_ad_full), AICc(lmer_ratiowt_ad_null)))

# Print the table with kable
anova_results_ratiowt_ad %>%
  kable(caption = "Results of comparison between full and null models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r, message=FALSE, warning=FALSE}
tab_model(lmer_ratiowt_ad_full, lmer_ratiowt_ad_null, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results, we select `1 + (1|Site) + (1|Sex) + (1|Year)` as the best-fitted model.
```{r}
shapiro.test(residuals(lmer_ratiowt_ad_null))
```

**Based on the measured variables, either without normal distribution of the residuals, we conclude that we did not identified a fixed effect influencing the body condition in adults, including the presence of *Babesia*.**\  

Visualisation: 
```{r}
plotG <- ggplot(data_ratiowt_ad_model, aes(x= Sex, y= RatioWT))+
  geom_point(aes(color=Year)) +
  geom_boxplot(alpha=0.5) +
  labs(x = "Sex",
       y = "Ratio mass/tarsus (body condition)", 
       title = "Body condition in adults according to sex, year and site")+ 
  theme(plot.title = element_text(size = 9)) + 
  facet_grid(~Site)

plotH <- ggplot(data_ratiowt_ad_model, aes(x= Status_qPCR, y= RatioWT, color=Status_qPCR, fill=Status_qPCR))+
  geom_point() +
  geom_boxplot(alpha=0.5) +
  labs(x = "Infection status",
       y = "Ratio mass/tarsus (body condition)", 
       title = "Body condition in adults according to sex, site, and infection status")+ 
  theme(plot.title = element_text(size = 9), 
        axis.text.x = element_blank()) + 
  facet_grid(Site~Sex)

ggarrange(plotG, plotH, nrow = 1, ncol = 2)
```

### 3.1.2. In chicks
We remove the missing values of variables that will be analysed in the model:
```{r}
data_ratiowt_ch_model <- data_ratiowt_chicks[!(is.na(data_ratiowt_chicks$Status_qPCR)), ]
data_ratiowt_ch_model <- data_ratiowt_ch_model[!(is.na(data_ratiowt_ch_model$Sex)), ]
```

We include the variables `Status_qPCR` as a fixed effect and `Site`, `Sex`, and `Year` as random effects in our full LMER model:
```{r, message=FALSE, warning=FALSE}
lmer_ratiowt_ch_full <- lmer(log_ratiowt ~ Status_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ch_model)
```

As the model is rather simple, we will directly use ANOVA and AICc comparison:
```{r, message=FALSE, warning=FALSE}
lmer_ratiowt_ch_null <- lmer(log_ratiowt ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ch_model)

anova6 <- anova(lmer_ratiowt_ch_full, lmer_ratiowt_ch_null, test="Chisq")

p_value_anova6 <- anova6$`Pr(>Chisq)`[2]

anova_results_ratiowt_ch <- data.frame(
  Comparison = c("Full",
                 "Full vs null"),
  ANOVA_p_value = c("-", p_value_anova6),
  AICc = c(AICc(lmer_ratiowt_ch_full), AICc(lmer_ratiowt_ch_null)))

# Print the table with kable
anova_results_ratiowt_ch %>%
  kable(caption = "Results of comparison between full and null models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r, message=FALSE, warning=FALSE}
tab_model(lmer_ratiowt_ch_full, lmer_ratiowt_ch_null, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results, we select `1 + (1|Site) + (1|Sex) + (1|Year)` as the best-fitted model.
```{r}
shapiro.test(residuals(lmer_ratiowt_ch_null))
```

**Based on the measured variables, we did not identified a fixed effect influencing the body condition in chucks, including the presence of *Babesia*.**\

Visualisation: 
```{r}
plotI <- ggplot(data_ratiowt_ch_model, aes(x= Sex, y= RatioWT, color=Sex, fill=Sex))+
  geom_point() +
  geom_boxplot(alpha=0.5) +
  labs(x = "Sex",
       y = "Ratio mass/tarsus (body condition)", 
       title = "Body condition in chicks according to sex and site")+ 
  theme(plot.title = element_text(size = 9), 
        axis.text.x = element_blank()) + 
  facet_grid(~Site)

plotJ <- ggplot(data_ratiowt_ch_model, aes(x= Status_qPCR, y= RatioWT, color=Status_qPCR, fill=Status_qPCR))+
  geom_point() +
  geom_boxplot(alpha=0.5) +
  labs(x = "Infection status",
       y = "Ratio mass/tarsus (body condition)", 
       title = "Body condition in chicks according to sex, site, and infection status")+ 
  theme(plot.title = element_text(size = 9), 
        axis.text.x = element_blank()) + 
  facet_grid(Site~Sex)

ggarrange(plotI, plotJ, nrow = 1, ncol = 2)
```


## 3.2. Based on parasitemia (continuous)
### 3.2.1. In adults
To determine whether a Pearson Correlation test can be used, the data must be tested for normality, autocorrelation and homoscedasticity:
```{r, message=FALSE, warning=FALSE}
shapiro.test(data_ratiowt_ad_ratioBab$RatioWT)
```
**p-value > 0.05, the data is normally distributed.**\

```{r, message=FALSE, warning=FALSE}
bptest(aov(RatioWT ~ Ratio_qPCR, data = data_ratiowt_ad_ratioBab)) 
```
**p-value > 0.05, the null hypothesis of residual homoscedasticity cannot be rejected.**\

```{r, message=FALSE, warning=FALSE}
dwtest(aov(RatioWT ~ Ratio_qPCR, data = data_ratiowt_ad_ratioBab))
```
**p-value > 0.05, the null hypothesis that the residuals are not autocorrelated cannot be rejected.**\

We can use a Pearson Correlation test: 
```{r, message=FALSE, warning=FALSE}
ratiowt_adults <- data_ratiowt_ad_ratioBab$RatioWT
ratioqPCR_adults <- data_ratiowt_ad_ratioBab$Ratio_qPCR
cor.test(ratiowt_adults,ratioqPCR_adults, method="pearson")
```

**p-value > 0.05 (p-value = 0.2357 & cor = 0.4128869), the null hypothesis that there is no correlation between the two variables `RatioWT` and `Ratio_qPCR` cannot be rejected.**\
**The parasitemia doesn't influence the body condition in adults.**


### 3.2.2. In chicks
We remove the missing values of variables that will be analysed in the model:
```{r}
data_ratiowt_ch_model <- data_ratiowt_chicks[!(is.na(data_ratiowt_chicks$Status_qPCR)), ]
data_ratiowt_ch_model <- data_ratiowt_ch_model[!(is.na(data_ratiowt_ch_model$Sex)), ]
data_ratiowt_ch_ratioBab <- data_ratiowt_ch_model[!(is.na(data_ratiowt_ch_model$Ratio_qPCR)), ]
data_ratiowt_ch_ratioBab <- data_ratiowt_ch_ratioBab[!(is.na(data_ratiowt_ch_ratioBab$ErrorRatio_qPCR)), ]
```

We include the variables `Ratio_qPCR` as a fixed effect and `Site`, `Sex`, and `Year` as random effects in our full LMER model. We also wanted to include `ErrorRatio_qPCR` and so we used a weighted LMER:
```{r, message=FALSE, warning=FALSE}
data_ratiowt_ch_ratioBab$log_ratio_qPCR <- log10(data_ratiowt_ch_ratioBab$Ratio_qPCR)
data_ratiowt_ch_ratioBab$weights <- 1 / data_ratiowt_ch_ratioBab$ErrorRatio_qPCR

lmer_ratiowt_ch_ratio_full <- lmer(RatioWT ~ log_ratio_qPCR + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ch_ratioBab)
```

As the model is rather simple, we will directly use ANOVA and AICc comparison:
```{r, message=FALSE, warning=FALSE}
lmer_ratiowt_ch_ratio_null <- lmer(RatioWT ~ 1 + (1|Site) + (1|Sex) + (1|Year), data=data_ratiowt_ch_ratioBab)

anova7 <- anova(lmer_ratiowt_ch_ratio_full, lmer_ratiowt_ch_ratio_null, test="Chisq")

p_value_anova7 <- anova7$`Pr(>Chisq)`[2]

anova_results_ratiowt_ratio_ch <- data.frame(
  Comparison = c("Full",
                 "Full vs null"),
  ANOVA_p_value = c("-", p_value_anova7),
  AICc = c(AICc(lmer_ratiowt_ch_ratio_full), AICc(lmer_ratiowt_ch_ratio_null)))

# Print the table with kable
anova_results_ratiowt_ratio_ch %>%
  kable(caption = "Results of comparison between full and null models") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We can also compare them with `tab_model` function:
```{r, message=FALSE, warning=FALSE}
tab_model(lmer_ratiowt_ch_ratio_full, lmer_ratiowt_ch_ratio_null, string.est = "IRR", show.aicc = TRUE, digits=2, show.ci = FALSE, show.intercept = FALSE)
```

Based on these results, the model with the lower AICc is `1 + (1|Site) + (1|Sex) + (1|Year)`.
```{r}
shapiro.test(residuals(lmer_ratiowt_ch_ratio_null))
```

**Based on the measured variables, we did not identified a fixed effect influencing the body condition in chucks, including the parasitemia.**\

Visualisation: 
```{r,message=FALSE, warning=FALSE}
ggplot(data_ratiowt_ch_ratioBab, aes(x = log_ratio_qPCR, y = RatioWT)) +
  geom_point() +  
  labs(x = "log10(Ratio_qPCR)",
       y = "Body condition", 
       title = "Body condition in chicks according to parasitemia") +  
  theme(plot.title = element_text(size = 12)) 
```


# 4. Summary
## 4.1. Circulating Hemoglobin Concentration
In adults, *Babesia* infection status does not affect hemoglobin concentration (LMER best-fitted model: `RatioWT + (1|Site) + (1|Sex) + (1|Year)`), nor does parasitemia (Pearson correlation, p = 0.3821).\ 

In chicks, *Babesia* infection status does not affect hemoglobin concentration (LMER best-fitted model: `1 + (1|Site) + (1|Sex) + (1|Year)`). The parasitemia tend to have an effect of concentration of hemoglobin in chicks, but the nature of this relationship is difficult to assess with our current data.

## 4.2. Body Condition
In adults, *Babesia* infection status does not affect body condition (LMER best-fitted model: `1 + (1|Site) + (1|Sex) + (1|Year)`), nor does parasitemia (Pearson correlation, p = 0.2357).\ 

In chicks, *Babesia* infection status does not affect body condition (LMER best-fitted model: `1 + (1|Site) + (1|Sex) + (1|Year)`), nor does the parasitemia.